The Design and Implementation of the Synchronous Language \CEU

### 2016-10-01 paper.tex
(>) syntax

In the subsections that follow, we discuss the main differences between \CEU and
Esterel:
%
Unique and queue-based external events
    (\ref{sec.ceu.exts});
Stack-based internal events
    (\ref{sec.ceu.ints});
Static concurrency checks
    (\ref{sec.ceu.checks});
Safe integration with \C
    (\ref{sec.ceu.c});
and
First-class synchronized timers
    (\ref{sec.ceu.timers}).
We finish the section with a summary of our design
    (\ref{sec.ceu.summary}).
%
We present the formal specification of the semantics of \CEU in a separate 
paper~\cite{ceu.phd}.

@@

In \CEU, internal events support stack-based micro reactions within external
reactions, providing more fine-grained control for intra-reaction execution.
%
Some variants of Statecharts~\cite{statecharts} also distinguish internal from
external events~\cite{statecharts.variants}.
%Some variants of Statecharts support local events which imply handling intra
%reactions like Céu.
In Statemate, ``reactions to external and internal events (...) can be sensed only
after completion of the step''~\cite{statecharts.statemate}, implying queue-based
execution.
In Stateflow, ``the receiving state (of the event) acts here as a
function''~\cite{statecharts.stateflow}, which is similar to Céu's stack-based
execution.
To avoid recursion, Céu adopts delayed awaits, while in Statemate, ``loops in
the broadcasting of events (between states) are forbidden''.

>>>>

Figure~\ref{lst.syn} presents the subset of the concrete syntax of \CEU used in
the paper as a quick reference.

\begin{figure}
\begin{lstlisting}
// DECLARATIONS

input  <type> <ids>;                        // external input  events
output <type> <ids>;                        // external output events
event  <type> <ids>;                        // internal events
var    <type> <id> = <exp>;                 // a variable with initial value

// EVENT HANDLING

<id> = await <id>;                          // await an event and assign received value
<id> = await <time>;                        // await time and assign delayed delta
emit <id> => <exp>;                         // emit an event passing a value

// CONTROL FLOW

<stmt> ; <stmt>                             // sequence
if <exp> then <stmts> else <stmts> end      // conditional
loop do <stmts> end                         // repetition
do <stmts> end                              // lexical block

par/or  do <stmts> with <stmts> end         // aborts when one side terminates
par/and do <stmts> with <stmts> end         // terminates when both sides terminate
par     do <stmts> with <stmts> end         // never terminates

finalize <stmts> with <stmts> end           // block finalization

// INTEGRATION WITH C

<_id>(<exps>)                               // C call (identifier starts with `_')
native do <stmts> end                       // declarations in C
native @const <_ids>                        // annotations: constant symbols
native @pure  <_ids>                        //              pure functions
native @safe  <_ids> with <_ids>            //              non-conflicting symbols
\end{lstlisting}
\caption{ Subset of the concrete syntax of \CEU used in the paper.
\label{lst.syn}
}
\end{figure}

In the subsections that follow, we discuss the main differences between \CEU and
Esterel:
%
Unique and queue-based external events
    (\ref{sec.ceu.exts});
Stack-based internal events
    (\ref{sec.ceu.ints});
Static concurrency checks
    (\ref{sec.ceu.checks});
Safe integration with \C
    (\ref{sec.ceu.c});
and
First-class synchronized timers
    (\ref{sec.ceu.timers}).
We finish the section with a summary of our design
    (\ref{sec.ceu.summary}).
%
We present the formal specification of the semantics of \CEU with a simplified
abstract syntax in a separate paper~\cite{ceu.phd}.

@@

In \CEU, internal events support stack-based micro reactions within external
reactions, providing more fine-grained control for intra-reaction execution.
%
Some variants of Statecharts~\cite{statecharts} also distinguish internal from
external events~\cite{statecharts.variants}.
%Some variants of Statecharts support local events which imply handling intra
%reactions like \CEU.
In Statemate, ``reactions to external and internal events (...) can be sensed only
after completion of the step''~\cite{statecharts.statemate}, implying queue-based
execution.
In Stateflow, ``the receiving state (of the event) acts here as a
function''~\cite{statecharts.stateflow}, which is similar to \CEU's stack-based
execution.
To avoid recursion, \CEU adopts delayed awaits, while in Statemate, ``loops in
the broadcasting of events (between states) are forbidden''.
